---
title: "Introduction to R"
author: "Women in Data Science"
output: 
  learnr::tutorial:
    css: "css/uoa.css"
    theme: "united"
    highlight: "pygments"
    progressive: true
    df_print: paged
runtime: shiny_prerendered
---

```{r setup, include = FALSE}
options(tibble.print_min = 5)
library(learnr)
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE, collapse = TRUE,
  fig.height = 4, fig.width = 8, fig.align = "center", cache = FALSE)
tutorial_html_dependency()

## need to set up any data frames we want to use here
```

## Kia ora koutou!

Welcome to this Introduction to R workshop: Adventu`R`es in `R`

The instructors are:

 + [Anna Fergusson](https://github.com/annafergusson)
 + [Earo Wang](https://github.com/earowang)
 + [Lisa Chen](https://unidirectory.auckland.ac.nz/profile/lisa-chen)
 + [Charlotte Jone-Todd](https://github.com/cmjt)
 + [Emma Lehrke](https://unidirectory.auckland.ac.nz/profile/emma-wilson)

## Data wrangling & visualisation

### To do

Chunk up the activities, have something at the beginning to recreate, for each chunk focus on one or two key ideas, just to get folks started!

### March madness!!

+ Animated gif showing March craziness on the trains, ferries and buses
+ Something to set the context for the exploration

https://at.govt.nz/about-us/reports-publications/at-metro-patronage-report/

### What is this plot and how can we recreate it?



### Packages

The main packages we'll be using for this part of the workshop are listed below.

```{r}
library(tidyverse)
library(lubridate)
library(glue)
library(janitor)
```

### Reading data into R

```{r read-data}
data_url <- "https://at.govt.nz/media/1982132/public-transport-patronage-data-january-2020.csv"
at_data_raw <- read_csv(data_url, skip = 4)
at_data_raw
```


### Quick look

```{r glimpse-data}
glimpse(at_data_raw)
```

### Easy cleaning


```{r clean-names}
at_data_raw %>% 
  rename_all(make_clean_names)
```

### Some `dplyr` functions

```{r}
at_data <- at_data_raw %>% 
  rename_all(make_clean_names) %>% 
  mutate(date = dmy(glue("01-{`month_year`}"))) %>%
  filter(year(date) >= 2006, year(date) < 2020) %>%
  select(date, total, bus = bus_total, train = train_rapid, ferry)
```

### And a plot!

```{r}
at_data %>% 
  ggplot(aes(x = date, y = total)) +
  geom_line() +
  geom_point()
```

### Nice plots

```{r}
at_data %>% 
  mutate(year = year(date), month = month(date, label = TRUE, abbr = TRUE)) %>% 
  ggplot(aes(x = month, y = total, group = year, colour = year)) +
  geom_line() +
  scale_colour_viridis_c()
```

### Even nicer plots

```{r}
at_data %>% 
  mutate(year = year(date), month = month(date, label = TRUE, abbr = TRUE)) %>% 
  ggplot(aes(x = year, y = total)) +
  geom_line() +
  facet_wrap(~ month)
```

### Reorganising/restructuring data

```{r}
at_data %>% 
  pivot_longer(cols = bus:ferry, 
               names_to = "types", 
               values_to = "tallies")
```

```{r}
at_data %>% 
  pivot_longer(cols = bus:ferry, 
               names_to = "service", 
               values_to = "travel") %>% 
  ggplot(aes(x = date, y = travel, colour = service)) +
  geom_line()
```

### Super plot fun times!

### Want to try later with more data?

Cycle count data here

https://at.govt.nz/cycling-walking/research-monitoring/monthly-cycle-monitoring/


## R & RStudio in the cloud

Put the link to the workspace/project here for particpants

Put some more info about how to set up R and RStudio themselves here for reference later


## Interactive spatial maps

Could just summarise what this part of the workshop did, with some reference links (but do this part in RStudio Cloud).

## Want to learn more?

Resources etc. for future learning :-)


